Learning filters from user provided image makers

### 2020-06-01 main.tex
add introduction

\usepackage[pdftex]{graphicx}
\ifCLASSOPTIONcompsoc
  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
\else
  \usepackage[caption=false,font=footnotesize]{subfig}
\fi
\usepackage{comment}
\usepackage{float}
\usepackage{mathtools}

\usepackage{subcaption}
\usepackage{siunitx}

@@

%%%%%%%%% TITLE
\title{Learning filters from markers}
\author{Italos Estilon de Souza\\
{\tt\small italos.souza@ic.unicamp.br}
% For a paper whose authors are all at the same institution,
% omit the following lines up until the closing ``}''.
% Additional authors and addresses can be added with ``\and'',
% just like the second author.
% To save space, use either the email address or home page, not both
\and
Alexandre Falcão\\
{\tt\small afalcao@ic.unicamp.br}
}

\maketitle
%\thispagestyle{empty}

%%%%%%%%% ABSTRACT

%%%%%%%%% BODY TEXT
\section{Introduction}

\section{Related Work}

\section{Preliminaries}
\begin{comment}
Let $D \subset \Z^d$ for $d > 1$ and let $\m{I}: D \to \R^b$, with $b \ge 1$, be a function that assings to each $p = (x_p, y_q) \in D$ a vector $\m{I}(p) = (I_1(p), I_2(p), \ldots, I_b(p))$.
A multi-band and multi-dimensional image is a pair ${\hat{I} = (D, \m{I})}$. We say that $D$ is the \textit{image domain}, $\m{I}(p)$ is the \textit{local feature vector} of a space element (\textit{spel}) $p \in D$, $b$ is the number of bands, and $d$ is the number of dimensions. In this work, we focus on images with $d=2$. The local feature vectors $\m{I}(p)$, of all $p\in D$, may be thought of as points of a \textit{local feature space} $\Re^b$.

An adjacency relation $\A \subseteq D \times D$ is a binary relation between spels $p, q \in D$. We focus on reflexive and translation-invariant relations. More precisely, we focus on adjacency relations given by \[\A_{i,j} = \{ (p, q) \in D : \|x_p - x_q\| \le i, \|y_p - y_q\| \le j \}.\] We say that $\A(p) = \{q \in D : (p, q) \in \A\}$ is the set of spels that are adjacent to $p$ with respect to $\A$.

Let $\D$ be a dataset of images that have the same image domain, and let $\lambda(\I) = \{\omega_1, \omega_2, \ldots, \omega_c\}$ be the true label of an image $\I \in \D$ among $c$ classes. A dataset $\D$ is \textit{supervised} if all image $\I \in \D$ is inspected by an expert and she agrees with $\lambda(\I)$. A dataset $\D$ is \textit{unsupervised} if $\lambda(\I) = \emptyset$ for all $\I \in \D$, that is all true labels are unkown. If there exist images with known and unkown true labels, we say that dataset $\D$ is \textit{semi-supervised}.

For a spel $p \in D$ and an adjacency relation $\A$, a \textit{locally expanded feature vector} is given by \[\E(p) = \lI(q_1) \oplus \lI(q_2) \oplus \ldots \oplus \lI(q_{|\A|})\] such as $q_i \in \A(p)$, for $i = 1, \ldots, |\A|$. Observe that for two different spels $p, q \in D$, it might be the case that $|\A(p)| \neq |\A(q)|$.

%resolver a definição de vetor extendido localmente para assegurar que todos os vetores tenham a mesma quantidade de dimensões

A multidimensioal and multi-band filter $\F$ is a pair $(\A, \mF)$ such that $\A \in D \times D$ is a adjacency relation and for all $(p, q) \in \A$, $\mF(p, q)$ assings a vector $(F_1(p, q), F_2(p, q), \ldots, F_b(p, q)) \in \R^b$ with $b \ge 1$ that satisfies the constraint that $\mF(p, q) = \mF(q - p)$. We can interpret $\F$ as a moving window that can be placed at any position $p \in D$. We say that $\bF = \{\F_1, \F_2, \ldots, \F_l\}$ is a \textit{filter bank} if for any $\F = (\A, \mF) \in \bF$ and $\F' = (\A', \mF') \in \bF$, $\A = \A'$, that is, all filters from $\bF$ has the same adjacency relation.

Now, we define the convolution operation. The convolution between an image $\I = (D, \lI)$ and a filter bank $\bF$ with $m$ filters produces a image $\J = (\pD, \lJ)$, in which, for $p \in \pD\subseteq D$, $\lJ(p) = (J_1(p), J_2(p), \ldots, J_m(p))$, and \[J_i(p) = \sum_{q \in \A(p)}\lI(q) \cdot {{\bf F}_i(q - p)},\] for $i = 1, 2,  \ldots, m$. For the sake of simplicity, we will assume that $\pD=D$.

In convolution neural networks, the filters has the purpose of detecting relevant features. This happens by producing an positive value for spels which have those features and negative values for the ones which does not have them. This information is handled by the activation function, which takes an image $\J = (D, \lJ)$ as input and produces another image $\hat{A} = (D, \m{A})$ as output.  In this work, we use the \textit{rectified linear unit}, which the elements of $\m{A}$ are defined as $A_i(p) = \max(0, J_i(p))$, $i=1,2,\ldots,m$, as activation function.

Pooling has the purpose of aggregating nearby information and makes convolutional neural networks less sensible to small variations in data. In this work, we use only max pooling which is an operation that takes an image $\J = (D, \lJ)$ with $m$ bands as input and produces an image $\hat{P} = (\pD, \m{P})$ where, for $p \in \pD\subseteq D$, ${\m{P}} (p) = (P_1(p), P_2(p), \ldots, P_m(p))$ such that \[P_i(p) = \max_{q \in \A(p)}\{I_i(q)\},\] for $i = 1, 2, \ldots, m$, and an adjacency relation $\A$. For the sake of simplicity, we will assume that $\pD=D$.

When computing the convolution between each image of a dataset $\D$ and each filter of a filter bank, followed by ReLU activation,
at each position $p\in D$, points $\E(p)$ from the
images $\I \in \D$ may be detected at the positive (or
negative) side of a hyperplane with normal vector $\mF_i$ at the origin of $\R^{b \times |\A|}$. The distance
between $\E(p)$ and that hyperplane is proportional to $J_i(p)
=\E(p) \cdot \eF_i$, for $i \in
\{1,2,\ldots,m\}$.

One can then imagine that whenever local patterns $\E(p)$ related to images from different categories appear as
distinct groups of points in $\R^{b\times |\A|}$, they should
be at different sides of the hyperplanes of the filter bank, so
different filters will be able to capture relevant details of distinct
classes. This is precisely the role of the bias in convolutional
neural networks. However, if the center of the points $\E(p)$ is at the origin of $\R^{b\times |\A|}$, the bias
computation becomes unnecessary.

Batch normalization is the operation that places the origin of
$\R^{b\times |\A|}$ at the center of the cloud of points
$\E(p)$, for a given $p\in D$, from all images $\hat{I}\in
\D$, and then makes those clouds of points for all $p\in D$ so
hyperspherical as possible \cite{ioffe2015batch}. This can be done independently, per
position and band, by transforming each image $\I=(D,\lI)$
into an image $\hat{N}=(D,{\bf N})$, ${\bf
  N}(p)=(N_1(p),N_2(p),\ldots,N_b(p))\in \R^b$, where, for $i\in \{1,2,\ldots,b\}$,
\begin{eqnarray}
  N_i(p) & = & \frac{I_i(p) - \mu_i(p)}{\sigma_i(p)} \\
  \mu_i(p) & = & \frac{1}{|\D|} \sum_{\forall \hat{I}\in \D} I_{i}(p) \\
  \sigma^2_i(p) & = & \frac{1}{|\D|} \sum_{\forall \hat{I}\in \D} \left(I_{i}(p)-\mu_i(p)\right)^2
\end{eqnarray}

A convolutional neural network may consist of a sequence of
convolutional layers followed by a pattern classifier. The
convolutional layers extract features from each input image $\I$
and create a global feature vector $\gN$ at the output of an
extra batch normalization after the last convolutional layer. The
classifier may be a linear Support Vector Machine (SVM) or a
Multi-Layer Perceptron (MLP). When using a MLP, the feature space of $\gN$ has usually
its dimensions reduced by a sequence of fully connected layers before
the decision layer.

For instance, assume that ${\gN}={\gN}^{(0)}\in \R^{m\times |D|}$
and let ${\W}^{(1)}_i \in \R^{m\times |D|}$ be the synaptic weight
vector of each neuron $i=1,2,\ldots,k_1$ in a first fully connected
layer with $k_1$ neurons. The result ${\gN}^{(1)}=(N^{(1)}_1,N^{(1)}_2,\ldots,N^{(1)}_{k_1})\in \R^{k_1}$, as
$k_1 = m\times |D|$, is obtained by $k_1$ inner products with ${\W}^{(1)}_i$, $i=1,2,\ldots,k_1$, followed by activation values, such that
\begin{eqnarray}
  N^{(1)}_i & = & \max\{0,\langle {\gN}^{(0)}, {\W}^{(1)}_i\rangle\}.
\end{eqnarray}

For a sequence of $l$ fully connected layers with $k_1,  k_2, \ldots, k_l$ neurons per layer, the above equation becomes recursive 
\begin{eqnarray}
  N^{(l)}_i & = & \max\{0,\langle {\gN}^{(l-1)}, {\W}^{(l)}_i\rangle\}, 
\end{eqnarray}
$i=1,2,\ldots,k_l$, creating a final feature vector ${\gN}^{(l)}\in
\R^{k_l}$ for classification by a decision layer.

For $c$ categories $\omega_1, \omega_2, \ldots, \omega_c$, the
decision layer consists of $c$ neurons with synaptic weights ${\W}_i\in \R^{k_l}$, $i=1,2,\ldots,c$, such that the input image
$\I$ is assigned to class $\omega_j$, $j\in \{1,2,\ldots,c\}$,
whenever
\begin{eqnarray}
  a_i & = & \max\{0,\langle {\gN}^{(L)}, {\W}_i\rangle\},\\ 
  a_j & = & \max_{i=1,2,\ldots,c}\{ a_i \}.
\end{eqnarray}

Given an image $\I = (D, \lI)$ of a dataset $\D$ such that each image has some category $\omega \in \{1, 2, \ldots, c\}$, a \textit{marker} is a pair $(p, \omega)$ where $p \in D$. We denote by $\M(\I)$ a set of markers for image $\I$. For a marker $m = (p, \omega) \in \M(\I)$, a \textit{patch} is the locally expanded feature vector $\E(p)$ of $p$ with respect to some adjacency relation $\A$. We denote by $P(\I, \M, \A)$ the set of patches of an image $\I$ with respect to a set of markers and an adjacency relation.
\end{comment}

@@

\begin{figure}
  \begin{center}
     \includegraphics[width=0.8\linewidth]{figures/filter.png}
  \end{center}
  \caption{Filter learned from clustering.}
  \label{fig:filter}
\end{figure}

@@

\begin{figure}
  \begin{center}
     \includegraphics[width=0.8\linewidth]{figures/tile0705-markers.png}
  \end{center}
     \caption{Example of markers.}
  \label{fig:ex-markers}
\end{figure}

To exemplify how our method works, we apply it to the problem of segmenting rock images into vegetation, rock and fractures. In Figure \ref{fig:ex-markers} we can see an example of the markers added by the network designer in regions that she considers relevant to characterize each of the classes.

\begin{figure*}
  \centering
  \begin{subfigure}[b]{0.33\textwidth}
      \includegraphics[width=\textwidth]{figures/probabilidade-fratura.png}
      \caption{Probability map for fractures.}
      \label{fig:prob-fractures}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{0.33\textwidth}
      \includegraphics[width=\textwidth]{figures/probabilidade-rocha.png}
      \caption{Probability map for rocks.}
    \label{fig:prob-rocks}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{0.33\textwidth}
      \includegraphics[width=\textwidth]{figures/probabilidade-vegetacao.png}
      \caption{Probability map for vegetation.}
  \label{fig:prob-vegetation}
  \end{subfigure}

  \caption{Probability maps for the classes.}
  \label{fig:prob_map}
\end{figure*}

We designed a neural network with two convolutional layers and used the features extracted by it to produce the probability maps that we can see in Figure \ref{fig:prob_map}, one for each class.

\begin{figure*}
  \centering
  \begin{subfigure}[b]{0.33\textwidth}
    \includegraphics[width=\textwidth]{figures/fratura.png}
    \caption{Segmentation of fractures.}
    \label{fig:segm-fractures}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{0.33\textwidth}
    \includegraphics[width=\textwidth]{figures/rocha.png}
    \caption{Segmentation of rocks.}
    \label{fig:segm-rocks}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{0.33\textwidth}
    \includegraphics[width=\textwidth]{figures/vegetacao.png}
    \caption{Segmentation of vegetation vegetation.}
  \label{fig:segm-vegetation}
\end{subfigure}

\caption{Segemetation of classes.}
\label{fig:segm}
\end{figure*}

The probability maps were used to produce the segmentations that we can see in Figure \ ref {fig: segm}. For simplify viewing, each image shows the segmentation of one class.

@@

@@

\begin{figure}
  \begin{center}
     \includegraphics[width=0.4\linewidth]{figures/coco_markers.png}
  \end{center}
     \caption{Markers on an image which contains a coconut tree.}
  \label{fig:coco-markers}
\end{figure}

\begin{figure}
  \begin{center}
     \includegraphics[width=0.4\linewidth]{figures/non_coco-markers.png}
  \end{center}
     \caption{Markers on an image which does not contain a coconut tree.}
  \label{fig:non-coco-markers}
\end{figure}

@@

\begin{table*}[h]
  \begin{center}
  \begin{tabular}{|l|c|c|c|}
  \hline
   Method & Precision & Recall & F-score \\
  \hline\hline
  Ours & $0.86 \pm 0.01$ & $0.84 \pm 0.02$ & $0.85 \pm 0.02$\\
  Ours (fine tuned) & $0.86 \pm 0.01$ & $0.84 \pm 0.02$ & $0.84 \pm 0.02$\\
  VGG-16 & $0.84 \pm 0.01$ & $0.76 \pm 0.03$ & $0.77 \pm 0.02 $ \\
  VGG-16 (fine tuned) & $0.84 \pm 0.01$ & $0.75 \pm 0.03$ & $0.77 \pm 0.02 $ \\
  \hline
  \end{tabular}
  \end{center}
  \caption{Results.}
  \label{tab:results}
  \end{table*}

\section{Conclusions and future work}

>>>>

\usepackage[pdftex]{graphicx}
\ifCLASSOPTIONcompsoc
  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
\else
  \usepackage[caption=false,font=footnotesize]{subfig}
\fi
\usepackage{comment}
\usepackage{float}
\usepackage{mathtools}
\usepackage{amsfonts}

@@

\title{Learning filters from user provided image makers}

% make the title area
\maketitle

\begin{abstract}

\end{abstract}

\section{Introduction}
Deep learning has proven to be applicable to different types of tasks, from image classification to the creation of synthetic data \cite{goodfellow2016deep}. Given the success of the application of deep learning in various areas of knowledge, this technique has also been applied to several tasks of remote sensing. Examples of this technique exist in various problems such as segmentation of terrain images \cite{kemker2018algorithms, kampffmeyer2016semantic, hamaguchi2018effective}, building identification \cite{xu2018building, lu2018detecting, liu2018multilevel}, and deforestation monitoring \cite{bragilevsky2017deep}.

The detection of trees is important to produce knowledge about the land use of a region, from the assessment of crop diversity or the performance of plantations in a given season to the assessment of damages after natural disasters such as drilling and flooding. As these plantations can span very large areas, manual detection can be costly and susceptible to errors.

In \cite{fassnacht2016review}, we can see a review of the problem of classification of tree species of aerial images. \cite{puttemans2018comparing} apply deep learning to the problem of detecting coconut trees in aerial images. Aparta et al. \cite{aparna2018cnn} apply convolutional neural networks to count coconut trees on plantations in India. Vargas-Muñoz et al. \cite{8899005} propose an active learning approach to detect coconut trees that seeks to reduce the effort of annotating images. The \cite{8899005} approach is divided into a stage of producing candidate patches for having a coconut tree and a stage where a CNN determines whether a patch has a coconut tree. We propose a different way to produce this classifier.

Despite this success, methods based on deep learning require a lot of annotated images for training, and annotating this volume of data is quite costly. Even interactive methods require large training sets. If these methods depend on huge training sets, their application may not be feasible in areas where it is impossible, or the cost of gathering and annotating these data sets is prohibitive. Much of the user's interaction with methods based on deep learning occurs through the annotation of the training set, and the user does not have much influence on the training process.

Despite the advances in deep learning over the last decades, especially with most recent deep neural networks, some important questions related to human interaction remain unanswered: (1) How to choose the most reasonable model, in the sense that it provides acceptable effectiveness, being as simple and efficient as needed, for a given classification problem? (2) How to train that model from a reduced number of samples?  (3) Can the user explain the decisions of that model? (4) Can that model improve from human corrections? The first question requires to explore human knowledge about the classification technique and the problem of interest. The second one requires to reduce as much as possible the human effort to train a model. The third issue is related to human understanding, and it may explore techniques (such as data visualization) to explain the decisions of the model and perhaps to guide human decisions concerning the design of the model. The fourth question is also important during training, and it is related to human control over the process. They all lead to the importance of involving human experts during the machine learning process.

In this paper, we present some advances towards the answers to the questions (1) and (2). We also discuss future research related to the questions (3) and (4). We present a method to the design of light-weighted convolutional neural networks from markers drawn by an expert on a few images to indicate representative regions of each object of interest for classification. From those markers, our approach can learn the filters of each layer, and the convolutional neural network can be designed in a layer-by-layer fashion, with no need for backpropagation in order to build a representative feature space for the classification problem. Backpropagation is only used to train a bank of multi-layer perceptrons for local pattern classification.

We propose a new way of learning the filters of the convolutional layers of deep neural networks that do not depend on the optimization of an error function and, therefore, do not need a large number of training samples. The training process is done with user intervention, which adds markers to the training images to indicate relevant features to characterize each class. To learn the filters, we get a patch around each marker of a class and cluster them. The centroid of each cluster is a filter that can be used to identify features of this cluster. The set of centroids of clusters of all clusters is the convolutional layer filters. For the next layer, the patches are generated on the output of the previous one. To evaluate our method, we use a dataset of aerial images to detect coconut trees. We identified that, with few layers and few training samples, we managed to get competitive results with state-of-the-art models.

@@

\begin{figure}[t]
  \begin{center}
     \includegraphics[width=0.8\linewidth]{figures/filter.png}
  \end{center}
  \caption{Filter learned from clustering.}
  \label{fig:filter}
\end{figure}

@@

To exemplify how our method works, we apply it to the problem of segmenting rock images into vegetation, rock and fractures. In Figure we can see an example of the markers added by the network designer in regions that she considers relevant to characterize each of the classes.

@@

\begin{figure}[t]
    \begin{center}
       \includegraphics[width=0.4\linewidth]{figures/coco_markers.png}
    \end{center}
       \caption{Markers on an image which contains a coconut tree.}
    \label{fig:coco-markers}
  \end{figure}

  \begin{figure}[t]
    \begin{center}
       \includegraphics[width=0.4\linewidth]{figures/non_coco-markers.png}
    \end{center}
       \caption{Markers on an image which does not contain a coconut tree.}
    \label{fig:non-coco-markers}
  \end{figure}

@@

@@

\begin{table}[!t]
  \begin{center}
  \begin{tabular}{|l|c|c|c|}
  \hline
   Method & Precision & Recall & F-score \\
  \hline\hline
  Ours & $0.86 \pm 0.01$ & $0.84 \pm 0.02$ & $0.85 \pm 0.02$\\
  Ours (fine tuned) & $0.86 \pm 0.01$ & $0.84 \pm 0.02$ & $0.84 \pm 0.02$\\
  VGG-16 & $0.84 \pm 0.01$ & $0.76 \pm 0.03$ & $0.77 \pm 0.02 $ \\
  VGG-16 (fine tuned) & $0.84 \pm 0.01$ & $0.75 \pm 0.03$ & $0.77 \pm 0.02 $ \\
  \hline
  \end{tabular}
  \end{center}
  \caption{Results.}
  \label{tab:results}
\end{table}
